---
title: "Standard Normal from Laplace distribution"
author: "Uche Mbaka"
date: "12/23/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

  The standard double exponential is symmetric exponential distribution, and since we know how to sample from exponential distribution using the inversion method, if we generate $Y \sim exp(1)$ we can generate $U(0,1)$ and sign $Y$, setting $X = |Y|$ if $U \leq 0.5$ and $X = -|y|$ if $U > 0.5$ where $X = \pm|X| \sim N(0,1)$.  
  
  $$Target \ distribution \ f(x)= (2\pi )^{\frac{-1}{2}}exp(\frac{-x^{2}}{2})$$
  $$Envelope function \ h(x)= \frac{1}{2}exp(-|x|)$$
  
  we set $w(x) = \frac{f(x)}{h(x)}$ and to minimize the probability of rejection we bound the ratio $\frac{f(x)}{h(x)}$ by the constant $M>0$ (ideally M should be close to 1). We set $g(x) = Mh(x)$. $M \geq Sup_{x}\{\frac{f(x)}{g(x)}\}$
  
  $$M = Max(w(x)) = w'(x) = 0$$
  $$w(x) = \frac{f(x)}{h(x)} = \frac{(2\pi )^{\frac{-1}{2}}exp(\frac{-x^{2}}{2})}{\frac{1}{2}exp(-|x|)}$$
  
  $$w(x) = exp(x- \frac{x^{2}}{2})\sqrt{\frac{2}{\pi}}$$
  
  The above result was gotten using **Mathematica**
  
  $$M = Max(w(x)) = Max(exp(x- \frac{x^{2}}{2})\sqrt{\frac{2}{\pi}})$$
  
  $$M = \sqrt{\frac{2e}{\pi}} \ at\  x = 1$$
  The above result was gotten from **Mathematica** using the Maximize function
  
  $$\frac{f(x)}{Mh(x)} = \frac{f(x)}{g(x)} = \frac{exp(x- \frac{x^{2}}{2})\sqrt{\frac{2}{\pi}}}{\sqrt{\frac{2e}{\pi}}}$$
  
  $$=exp(-\frac{1}{2}+x+\frac{x^{2}}{2})$$
  
  Above result gotten from **Mathematica**
  
  $$\frac{f(x)}{g(x)} = exp(-\frac{1}{2}+x+\frac{x^{2}}{2}) = exp(-\frac{(x-1)^{2}}{2})$$
  
  The algorithm for generating $X \sim N(0,1)$  
    1. Generate $Y \sim Exp(1)$ using inversion method i.e ($Y = -log(1-U); \ U\sim U(0,1)$)  
    2. Generate $U \sim U(0,1)$  
    3. If $U \leq exp(-\frac{(y-1)^{2}}{2})$, set $|X| = Y$; else go back to 1.
    4. Generate $U \sim U(0,1)$, set $X = |X|$ if $U \leq 0.5$, else set $X = -|X| \ (U > 0.5)$
    
 ```{r}
    stdNormal <- function(n){
      X <- vector(length = n)
      for(i in 1:n){
        while(TRUE){
          y <- -log(1-runif(1))#exp(1)
          u <- runif(1)
          if(u <= exp(-((y - 1)^2/2))){
            u2 <- runif(1)
            if(u2 <= 0.5){#sign Y
              X[i] <- y 
              break
            }else{
              X[i] <- -y
              break
            }#inner if else() 
          }#outer if()
        }#while()
      }#for()
      return (X)
    }#stdNormal
    
    #Check
    X <- stdNormal(1000)
    
    hist(X, freq = FALSE, main = "Density Plot of X")

    lines(density(X, adjust=2), lwd = 2, col = 2)
    
  ```